---
title: "Analysis of PBMCs from individuals with severe COVID-19"
author: "Jose Alquicira Hernandez"
date: "30/09/2020"
output: 
  html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)
```

# Import libraries

Install packages if necessary

```{r installation}
if (!requireNamespace("tidyverse", quietly = TRUE))
    install.packages("tidyverse")
if (!requireNamespace("here", quietly = TRUE))
    install.packages("here")
if (!requireNamespace("Seurat", quietly = TRUE))
    install.packages("Seurat")
```

Load libraries

```{r load_libraries}
library("tidyverse")
library("here")
library("Seurat")
library("patchwork")
```

Set up colors to match populations with publication

```{r colors}
pal <- c("#1F77B4", "#AEC7E8", "#FF7F0E", "#FFBB78", "#2CA02C",
         "#98DF8A", "#D62728", "#FF9896", "#9467BD", "#C5B0D5",
         "#8C564B", "#C49C94", "#E377C2", "#F7B6D2", "#7F7F7F",
         "#C7C7C7", "#BCBD22", "#DBDB8D", "#17BECF", "#9EDAE5")
```

# Import data

Read data

```{r read_data}
filename <- here("data", "blish_covid.seu.rds")
data <- readRDS(filename)
```


# Label cells according to publication figures

```{r label_cells}
x <- data@meta.data$cell.type 
data$cell_type <- case_when(
  x == "Activated Granulocyte" ~ "Developing neutrophil",
  x == "CD4 T"                 ~ "IFN-stim CD4 T",
  x == "CD8eff T"              ~ "Proliferative lymphocytes",
  x == "Class-switched B"      ~ "IgM PB",
  TRUE ~ x
)
Idents(data) <- "cell_type"
```

# Reproduce original figure from paper

```{r original_fig}
DimPlot(data, label = TRUE, label.size = 3, repel = TRUE, cols = pal)
```

In the paper, the following variables in the metadata slot are regressed out
within `SCTransform`.

```{r covariates}
data@commands$SCTransform.RNA$vars.to.regress
```

And the command run was

```{r command}
data@commands$SCTransform.RNA
```

For reproducibility purposes, let's use the original command as reference
`return.only.var.genes` is set to `TRUE` by default so 
`return.only.var.genes = TRUE` will be ommitted.

For the UMAP and to construct the Shared Nearest Neighbor (SNN) Graph, *50* 
principal components were used:


```{r dims}
length(data@commands$RunUMAP.SCT.pca$dims)
length(data@commands$FindNeighbors.SCT.pca$dims)
```

As we're only interested in the scaled pearson residuals, let's set
`do.correct.umi = FALSE`.


# Regress out mitochondrial and ribosomal gene expression only

Set up future settings for parallelization

```{r future_settings}
options(future.globals.maxSize = 1.5 * 1024^3)
future::plan("multisession", workers = 2)
```


```{r sct_1, message=FALSE, warning=FALSE, results='hide'}
data <- SCTransform(data, vars.to.regress = c("percent.mt", 
                                              "percent.rps", 
                                              "percent.rpl", 
                                              "percent.rrna"), 
                    do.correct.umi = FALSE)

data <- RunPCA(data)
data <- RunUMAP(data, dims = 1:50)
```

```{r plot_1}
(p <- DimPlot(data, label = TRUE, label.size = 3, repel = TRUE, cols = pal))
```

```{r save_plot_1}
ggsave(here("mito_and_ribo.png"), p, width = 8, height = 5, dpi = "print")
```


# Regress out mitochondrial gene expression only


```{r sct_2, message=FALSE, warning=FALSE, results='hide'}
data <- SCTransform(data, vars.to.regress = "percent.mt", 
                    do.correct.umi = FALSE)


data <- RunPCA(data)
data <- RunUMAP(data, dims = 1:50)
```

```{r plot_2}
(p2 <- DimPlot(data, label = TRUE, label.size = 3, repel = TRUE, cols = pal))
```

```{r save_plot_2}
ggsave(here("mito.png"), p, width = 8, height = 5, dpi = "print")
```


# Subset neutrophils, dev. neutruphils, B cells, and plasmablasts


```{r sct_3, message=FALSE, warning=FALSE, results='hide'}
i <- data$cell_type %in% c("Developing neutrophil", "IgM PB", "IgA PB", "IgG PB", "Neutrophil", "B")
data <- data[,i]

Idents(data) <- "cell_type"

data <- SCTransform(data,  vars.to.regress = c("percent.mt", "percent.rps", "percent.rpl",
                                               "percent.rrna", "nCount_RNA", "nFeature_RNA"),
                    do.correct.umi = FALSE)


data <- RunPCA(data)
data <- RunUMAP(data, dims = 1:50)
```


```{r plot_3}
cols2 <- c("#AEC7E8", "#FF7F0E", "#FFBB78", "#9467BD", "#E377C2", "#DBDB8D")
(p <- DimPlot(data, label = TRUE, label.size = 3, repel = TRUE, cols = cols2))
```

```{r save_plot_3}
ggsave(here("neutrophils_b_plasma.png"), p, width = 8, height = 5, dpi = "print")
```


# Session information

```{r session_info}
options(width = 120)
devtools::session_info()
```

